---
title: "R Notebook"
output: html_notebook
---

# Training a neural net. Basic structure and setup.
```{r}
# Create list with required libraries
libraries <- c("dplyr", "readr", "reticulate", "ggplot2", "Metrics", "tidyverse", "glmnet")


# Loop through each library and check if it's installed, if not, install it and then load it to include in this project
for (lib in libraries)
{
  if (!require(lib, character.only = TRUE))
  {
    install.packages(lib)
    library(lib, character.only = TRUE)
  }
}

```

# Data Import
```{r}
# Storing Data sets as data frames

revenue <- read.csv('./data/original/umsatzdaten_gekuerzt.csv')
kiwo <- read.csv('./data/original/kiwo.csv')
ferien <- read.csv('./data/original/schulferienSH.csv')
feiertage <- read.csv('./data/original/feiertage.csv')
weather <- read.csv('./data/original/weather.csv')

```

# Joining the kiwo data with the revenue data
## Replacing all instances of Kiwo = NA with Kiwo = 0
## Reformating Column KielerWoche to boolean values

```{r}
df <- full_join(revenue, kiwo, by = "Datum") %>%
  replace_na(list(KielerWoche = 0)) %>%
  mutate(KielerWoche = ifelse(KielerWoche == 1, TRUE, FALSE))

```

# Joining the feiertage data with the revenue data
## Replacing all instances of feiertage = NA with feiertage = 0
## Reformating Column Feiertage to boolean values

# Assuming df is your data frame and "Datum" is the name of the column containing date strings

```{r}
# Load the feiertage data
feiertage <- read.csv('./data/original/feiertage.csv')

# Convert the "Datum" column to a Date
feiertage$Datum <- as.Date(feiertage$Datum, format = "%Y-%m-%d")

# View the feiertage data
head(feiertage)
````

```{r}
df <- full_join(revenue, feiertage, by = "Datum") %>%
  replace_na(list(feiertage = 0)) %>%
  mutate(feiertage = ifelse(feiertage == 1, TRUE, FALSE))

```

# Joining the weather data with the df containing revenue and kiwo

```{r}
df <- full_join(df, weather, by = "Datum")

```

# Cleaning up the revenue data to convert Datum column to correct data type (i.e. char -> date)
```{r}
df$Datum <- as.Date(df$Datum, format = "%Y-%m-%d")
```


# Converting the  "Warengruppe" to char, as the algorithm might interpret the numeric values as ordered, while they are categorial.
```{r}
df <- df %>%
  mutate(Warengruppe = ifelse(Warengruppe == 1, "Brot", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 2, "Brötchen", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 3, "Croissant", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 4, "Konditorei", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 5, "Kuchen", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 6, "Saisonbrot", Warengruppe))
```

# Adding weekday column to data set
```{r}
df$Wochentag <- weekdays(df$Datum)

```

# Adding holiday column to data set
```{r}
ferien$Datum <- as.Date(ferien$Datum, format = "%Y-%m-%d")

df <- full_join(df, ferien, by = "Datum") %>%
  replace_na(list(FerienSH = 0)) %>%
  mutate(FerienSH = ifelse(FerienSH == 1, TRUE, FALSE))

```

# Adding feiertage column to data set
```{r}
feiertage$Datum <- as.Date(feiertage$Datum, format = "%Y-%m-%d")

df <- full_join(df, feiertage, by = "Datum") %>%
  replace_na(list(feiertage = 0)) %>%
  mutate(feiertage = ifelse(feiertage == 1, TRUE, FALSE))

```

# This helps us figure out if we have NA values in the dataset
```{r}
any(is.na(df))
```
# Deleting all rows that contain even one NA value (This section will be replaced with data imputation later on.)
```{r}
# Check for complete cases
complete_rows <- complete.cases(df)

# Subset the data frame to keep only complete rows
clean_df <- df[complete_rows, ]

#From here on out I'll work with the clean_df data set. Once the imputation step is implemented, the following code can be used for both.
```

# Simple Linear regression model:

#### Teilen des Datensatzes in Trainings- Validierungs- und Testdatensatz
```{r}
# Set a random seed for reproducibility
set.seed(42)

# Shuffle the data
data_shuffled <- clean_df %>% 
  sample_frac(1)

# Calculate the number of rows for each dataset
n_total <- nrow(clean_df)
n_train <- floor(0.7 * n_total)
n_validation <- floor(0.20 * n_total)

# Split the data into training, validation, and test datasets
train_data <- data_shuffled %>%
  slice(1:n_train)

validation_data <- data_shuffled %>%
  slice((n_train + 1):(n_train + n_validation))

test_data <- data_shuffled %>%
  slice((n_train + n_validation + 1):n_total)

# Check the dimensions of the datasets
cat("Training dataset dimensions:", dim(train_data), "\n")
cat("Validation dataset dimensions:", dim(validation_data), "\n")
cat("Test dataset dimensions:", dim(test_data), "\n")

```



### Beispiel einer einfachen linearen Regression
```{r}
  mod <- lm(Umsatz ~ Datum, clean_df)
    summary(mod)
  mod_1 <- lm(Umsatz ~ Datum + as.factor(Warengruppe), clean_df)
    summary(mod_1)
  mod_2 <- lm(Umsatz ~ Datum + as.factor(Warengruppe) + as.factor(Wochentag), clean_df)
    summary(mod_2)
  mod_3 <- lm(Umsatz ~ Datum + as.factor(Warengruppe) + as.factor(Wochentag) + FerienSH, clean_df)
    summary(mod_3)
  mod_4 <- lm(Umsatz ~ Datum + as.factor(Warengruppe) + as.factor(Wochentag) + FerienSH + Temperatur, clean_df)
    summary(mod_4)
  mod_5 <- lm(Umsatz ~ Datum + as.factor(Warengruppe) + as.factor(Wochentag) + FerienSH + Temperatur + KielerWoche, clean_df)
    summary(mod_5)
  mod_6 <- lm(Umsatz ~ Datum + as.factor(Warengruppe) + as.factor(Wochentag) + FerienSH + Temperatur + KielerWoche + Bewoelkung, clean_df)
    summary(mod_6)
    # Quick comment: KiWo and Bewoelkung do not increase R-squared
    mod_7 <- lm(Umsatz ~ Datum + as.factor(Warengruppe) + as.factor(Wochentag) + FerienSH + Temperatur + KielerWoche + Bewoelkung + Windgeschwindigkeit, clean_df)
    summary(mod_7)
    # Windgeschwindigkeit only inproves R-squared marginally.
  mod_8 <- lm(Umsatz ~ Datum + as.factor(Warengruppe) + as.factor(Wochentag) + FerienSH + Temperatur + KielerWoche + Bewoelkung + Windgeschwindigkeit + Wettercode, clean_df)
    summary(mod_8)
  
  
```

### Nutzung des resultierenden Modells für eine Vohersage
```{r}
# Make predictions using the test data
predicted_values <- predict(mod, newdata = validation_data)

# Compare the predicted values with the actual values
comparison <- data.frame(Actual = validation_data$Umsatz, Predicted = predicted_values)

# Calculate the mean squared error (RMSE)
rmse <- sqrt(mean((comparison$Actual - comparison$Predicted)^2))

# Display the comparison and RMSE
head(comparison)
cat("Root Mean Squared Error (RMSE):", rmse, "\n")

```


### Beispiel einer linearen Regression mit Regularisierung
```{r}
mod <- glmnet(as.matrix(train_data[c('Datum', 'Temperatur')]), train_data$Umsatz)
mod

```

