---
title: "R Notebook"
output: html_notebook
---

# Training a neural net. Basic structure and setup.
```{r}
# Create list with required libraries
libraries <- c("dplyr", "readr", "reticulate", "ggplot2", "Metrics", "tidyverse")


# Loop through each library and check if it's installed, if not, install it and then load it to include in this project
for (lib in libraries)
{
  if (!require(lib, character.only = TRUE))
  {
    install.packages(lib)
    library(lib, character.only = TRUE)
  }
}

```

# Data Import
```{r}
# Storing Data sets as data frames

revenue <- read.csv('./data/original/umsatzdaten_gekuerzt.csv')
kiwo <- read.csv('./data/original/kiwo.csv')
ferien <- read.csv('./data/original/schulferienSH.csv')
weather <- read.csv('./data/original/weather.csv')

```


# Joining the kiwo data with the revenue data
## Replacing all instances of Kiwo = NA with Kiwo = 0
## Reformating Column KielerWoche to boolean values

```{r}
df <- full_join(revenue, kiwo, by = "Datum") %>%
  replace_na(list(KielerWoche = 0)) %>%
  mutate(KielerWoche = ifelse(KielerWoche == 1, TRUE, FALSE))

```


# Joining the weather data with the df containing revenue and kiwo

```{r}
df <- full_join(df, weather, by = "Datum")

```

# Cleaning up the revenue data to convert Datum column to correct data type (i.e. char -> date)
```{r}
df$Datum <- as.Date(df$Datum, format = "%Y-%m-%d")
```


# Converting the  "Warengruppe" to char, as the algorithm might interpret the numeric values as ordered, while they are categorial.
```{r}
df <- df %>%
  mutate(Warengruppe = ifelse(Warengruppe == 1, "Brot", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 2, "Brötchen", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 3, "Croissant", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 4, "Konditorei", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 5, "Kuchen", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 6, "Saisonbrot", Warengruppe))
```

# Adding weekday column to data set
```{r}
df$weekday <- weekdays(df$Datum)

```

# Adding holiday column to data set
```{r}
ferien$Datum <- as.Date(ferien$Datum, format = "%Y-%m-%d")

df <- full_join(df, ferien, by = "Datum") %>%
  replace_na(list(FerienSH = 0)) %>%
  mutate(FerienSH = ifelse(FerienSH == 1, TRUE, FALSE))

```


# This helps us figure out if we have NA values in the dataset
```{r}
any(is.na(df))
```
# Deleting all rows that contain even one NA value (This section will be replaced with data imputation later on.)
```{r}
# Check for complete cases
complete_rows <- complete.cases(df)

# Subset the data frame to keep only complete rows
clean_df <- df[complete_rows, ]

#From here on out I'll work with the clean_df data set. Once the imputation step is implemented, the following code can be used for both.
```





