---
title: "dataPreparationPhilipp"
author: "Philipp Meisinger, Olav Cornelius, Alexander Ruoff, Felix Wehkamp"
date: "`r Sys.Date()`"
output: html_document
---

## Datenaufbereitung und Visualisierung
We have to install the Python environment in order to run the following pieces of code.
```{r message=FALSE, warning=FALSE, include=FALSE}
# Create list with required libraries
libraries <- c("dplyr", "readr", "ggplot2", "tidyverse", "lubridate", "VIM")


# Loop through each library and check if it's installed, if not, install it and then load it to include in this project
for (lib in libraries)
{
  if (!require(lib, character.only = TRUE))
  {
    install.packages(lib)
    library(lib, character.only = TRUE)
  }
}

```

### Data Import

```{r include=FALSE}
# Storing Data sets as data frames

revenue <- read.csv('../data/umsatzdaten_gekuerzt.csv')
kiwo <- read.csv('../data/kiwo.csv')
ferien <- read.csv('../data/schulferienSH.csv')
weather <- read.csv('../data/weather.csv')

```

### Joining the kiwo data with the revenue data

Replacing all instances of Kiwo = NA with Kiwo = 0 Reformating Column KielerWoche to boolean values

```{r include=FALSE}
df_prep <- left_join(revenue, kiwo, by = "Datum") %>%
  replace_na(list(KielerWoche = 0)) %>%
  mutate(KielerWoche = ifelse(KielerWoche == 1, TRUE, FALSE))

```

### Joining the weather data with the df containing revenue and kiwo

```{r include=FALSE}
df_prep <- left_join(df_prep, weather, by = "Datum")

```

### Cleaning up the revenue data to convert Datum column to correct data type (i.e. char -\> date)

```{r include=FALSE}
df_prep$Datum <- as.Date(df_prep$Datum, format = "%Y-%m-%d")
```

### Converting the "Warengruppe" to char, as the algorithm might interpret the numeric values as ordered, while they are categorial.

```{r include=FALSE}
df_prep <- df_prep %>%
  mutate(Warengruppe = ifelse(Warengruppe == 1, "Brot", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 2, "Brötchen", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 3, "Croissant", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 4, "Konditorei", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 5, "Kuchen", Warengruppe)) %>%
  mutate(Warengruppe = ifelse(Warengruppe == 6, "Saisonbrot", Warengruppe))
```

### Adding weekday column to data set

```{r include=FALSE}
df_prep$Wochentag <- weekdays(df_prep$Datum)
```

### Adding holiday column to data set

```{r include=FALSE}
ferien$Datum <- as.Date(ferien$Datum, format = "%Y-%m-%d")

df_prep <- left_join(df_prep, ferien, by = "Datum") %>%
  replace_na(list(FerienSH = 0)) %>%
  mutate(FerienSH = ifelse(FerienSH == 1, TRUE, FALSE))
```

### Adding Month variable
```{r}
df_prep$Monat <- month(df_prep$Datum)
```


### This helps us figure out if we have NA values in the dataset

```{r include=FALSE}
any(is.na(df_prep))

any(is.na(df_prep$Bewoelkung))

any(is.na(df_prep$Temperatur))
any(is.na(df_prep$Windgeschwindigkeit))
any(is.na(df_prep$Wettercode))

missing_values <- df_prep %>%
  filter(is.na(Temperatur) & is.na(Bewoelkung) & is.na(Windgeschwindigkeit) & is.na(Wettercode))

```

### Data Imputation
```{r}

# VIM Aggregation Plot
df_prep %>%
  aggr(combined=TRUE, numbers=TRUE)

# Listwise deletion
# df_deletion <- na.omit(df_prep)
# df_deletion %>%
#  aggr(combined=TRUE, numbers=TRUE)

# VIM Hot-Deck Imputation
df_hotdeck1 <- df_prep %>%  hotdeck()
df_hotdeck1 %>%
  aggr(combined=TRUE, numbers=TRUE)

 # ggplot(df_hotdeck1) +
 #  geom_point(aes(
 #    x=Temperatur,
 #    y=Umsatz,
 #    color=Temperatur_imp
 #    )
 #  )

df_hotdeck2 <- df_prep %>% 
  hotdeck(ord_var = "Datum")

df_hotdeck2 %>%
  aggr(combined=TRUE, numbers=TRUE)

# ggplot(df_hotdeck2) +
#  geom_point(aes(
#    x=Bewoelkung,
#    y=Temperatur,
#    color=Temperatur_imp
#    )
#  )

sleep_hotdeck3 <- sleep_hotdeck2 %>%  hotdeck(ord_var = "Sleep")
sleep_hotdeck3 %>% aggr(combined=TRUE, numbers=TRUE)

# VIM Iterative Robust Model Imputation (IRMI)
sleep_irmi <- sleep %>%  irmi()
ggplot(sleep_irmi) +
  geom_point(aes(x=Sleep, y=Dream, color=Sleep_imp))


# Impact
# cor(sleep_deletion$Sleep, sleep_deletion$Dream)
cor(df_hotdeck1$Datum, sleep_hotdeck1$Temperatur, use = "complete.obs")
cor(df_hotdeck2$Sleep, sleep_hotdeck2$Temperatur, use = "complete.obs")
cor(df_hotdeck3$Sleep, sleep_hotdeck3$Temperatur, use = "complete.obs")
# cor(df_irmi$Sleep, sleep_irmi$Dream)
```



### Deleting all rows that contain even one NA value (This section will be replaced with data imputation later on.)

```{r include=FALSE}
# Check for complete cases
# complete_rows <- complete.cases(df_prep)

# Subset the data frame to keep only complete rows
# clean_df <- df_prep[complete_rows, ]

# write_csv(clean_df, "../data/data.csv")

#From here on out I'll work with the clean_df data set. Once the imputation step is implemented, the following code can be used for both.
```