---
title: "dataPrep"
author: "Philipp Meisinger, Olav Cornelius, Alexander Ruoff, Felix Wehkamp"
date: "`r Sys.Date()`"
output: html_document
---

## Datenaufbereitung und Visualisierung
```{r message=FALSE, warning=FALSE, include=FALSE}
# Create list with required libraries
libraries <- c("dplyr", "readr", "ggplot2", "tidyverse", "lubridate", "VIM")


# Loop through each library and check if it's installed, if not, install it and then load it to include in this project
for (lib in libraries)
{
  if (!require(lib, character.only = TRUE))
  {
    install.packages(lib)
    library(lib, character.only = TRUE)
  }
}

remove(libraries)

```

### Data Import

```{r include=FALSE}
# Storing Data sets as data frames

revenue <- read.csv('../data/umsatzdaten_gekuerzt.csv')
kiwo <- read.csv('../data/kiwo.csv')
ferien <- read.csv('../data/schulferienSH.csv')
weather <- read.csv('../data/weather.csv')
prediction_2019 <- read.csv('../data/prediction_template.csv')

```


### Joining in the Test data set.
The way we are to predict the "Umsatz" values for the times between the 2019-06-08 and the 2019-07-30 is by filling out the values we know in that time frame as well. So we'll include this time frame in the data preparation step, so that we can later on split the data set and keep this one as the test data set.

```{r include=FALSE}
df_prep <- full_join(revenue, prediction_2019, by = "Datum") %>%
  mutate(Warengruppe.x = ifelse(is.na(Warengruppe.x), Warengruppe.y, Warengruppe.x)) %>%
  mutate(Umsatz.x = ifelse(is.na(Umsatz.x), Umsatz.y, Umsatz.x))

df_prep <- df_prep %>% 
  rename(Umsatz = Umsatz.x, Warengruppe = Warengruppe.x)

df_prep <- select(df_prep, -Warengruppe.y)
df_prep <- select(df_prep, -Umsatz.y)

```

### Joining the kiwo data

Replacing all instances of Kiwo = NA with Kiwo = 0 Reformating Column KielerWoche to boolean values the weather data with the df containing revenue and kiwo

```{r include=FALSE}
df_prep <- left_join(df_prep, kiwo, by = "Datum") %>%
  replace_na(list(KielerWoche = 0)) %>%
  mutate(KielerWoche = ifelse(KielerWoche == 1, TRUE, FALSE))

```


```{r include=FALSE}
df_prep <- left_join(df_prep, weather, by = "Datum")

```


```{r}
# Deleting Column Wettercode from df:
df_prep <- select(df_prep, -Wettercode)

```

### Cleaning up the revenue data to convert Datum column to correct data type (i.e. char -\> date)

```{r include=FALSE}
df_prep$Datum <- as.Date(df_prep$Datum, format = "%Y-%m-%d")

```

### Adding weekday column to data set

```{r include=FALSE}
df_prep$Wochentag <- weekdays(df_prep$Datum)
```

### Adding holiday column to data set

```{r include=FALSE}
ferien$Datum <- as.Date(ferien$Datum, format = "%Y-%m-%d")

df_prep <- left_join(df_prep, ferien, by = "Datum") %>%
  replace_na(list(FerienSH = 0)) %>%
  mutate(FerienSH = ifelse(FerienSH == 1, TRUE, FALSE))
```

### Adding Month variable
```{r}
df_prep$Monat <- month(df_prep$Datum)


# df_prep <- df_prep[df_prep$Monat == 6 | df_prep$Monat == 7, ]
# df
```


```{r}

df_prep_lag <- df_prep %>%
  arrange(Datum, Warengruppe) %>%
  group_by(Warengruppe) %>%
  mutate(value_prev_year = lag(Umsatz, default = NA, n = 365)) %>%
  ungroup()



df_prep_lag$value_prev_year <- ifelse(is.na(df_prep_lag$value_prev_year), df_prep_lag$Umsatz, df_prep_lag$value_prev_year)


```


### Data Imputation
We first preview the missing data correlations and then use a hotdeck imputation to fill in the missing values.
```{r}

# VIM Aggregation Plot
df_prep %>%
  aggr(combined=TRUE, numbers=TRUE)

# Hotdeck imputation with VIM visualization for data set df_prep
df_hotdeck_1 <- df_prep %>%  hotdeck()

df_hotdeck_1 %>%
  aggr(combined=TRUE, numbers=TRUE)

df_hotdeck_1_clean <- df_hotdeck_1 %>%
  select(-ends_with('_imp'))


# Hotdeck imputation with VIM visualization for data set df_prep_lag
df_hotdeck_1_lag <- df_prep_lag %>% hotdeck()

df_hotdeck_1_clean_lag <- df_hotdeck_1_lag %>%
  select(-ends_with('_imp'))

```

Making sure the data set is sorted by date
```{r}
df_hotdeck_1 <- df_hotdeck_1 %>%
  arrange(Datum)
df_hotdeck_1_clean <- df_hotdeck_1_clean %>%
  arrange(Datum)

```

Saving the Data sets into .csv files.
```{r include=FALSE}

write_csv(df_hotdeck_1_clean, "../data/data_clean.csv")
write_csv(df_hotdeck_1_clean_lag, "../data/data_clean_lag.csv")

```